Factoring RSA keys with Shor's algorithm

### 2020-06-02 paper/paper.tex
paper: Add mod exp + misc

\subsection{Order-finding subroutine}\label{order-finding}
We can further reduce the order-finding problem to the well-known problem of finding the eigenvalues of a unitary operator, which can be solved with the quantum phase estimation algorithm. Shor exploits the unitary operator
\begin{equation}
U_a: x \rightarrow ax \mod N
\end{equation}
whose eigenvalues\footnote{The reader can prove this by substitution: the eigenvectors take the form $\frac{1}{\sqrt{r}}\sum_{m=0}^{r-1}\exp(\frac{-2\pi i m n}{r})\ket{x^k \mod N}$.} are a function of $r$:
\begin{equation}
e^{2i\pi k/r}\text{ for some integer }k
\end{equation}

@@

@@

\subsection{Application}
We develop a simple client-server application that simulates an email service. Users can log in with their username and password, and once logged in they can browse their email messages. It uses a simple cryptographic protocol that simulates TLS, where the server sends its public key to the client, which in turn sends an encrypted AES key (a "session key") allowing for secure two-way communication. The goal of this demonstration is to capture the public key and factor it, allowing a third party to decrypt messages intended for the server and snoop on the credentials and the email messages.

@@

\begin{thebibliography}{00}
\bibitem{shor} P. W. Shor, "Algorithms for quantum computation: discrete logarithms and factoring", SIAM J.Sci.Statist.Comput. 26, 1997
\bibitem{rsa} R. Rivest, A. Shamir, L. Adleman, "A Method for Obtaining Digital Signatures and Public-Key Cryptosystems", Communications of the ACM 21, 1978
\bibitem{miller} G. L. Miller, "Riemann's hypothesis and tests for primality", J. Comput. System Sci., 1976
\bibitem{vbe} V. Vedral, A. Barenco, A. Ekert, "Quantum Networks for Elementary Arithmetic Operations", Physical Review A 54
\bibitem{ibm} IBM, "Shor's algorithm",\\https://quantum-computing.ibm.com/docs/guide/q-algos/shor-s-algorithm
\bibitem{github-repo} G. Muscarello, https://github.com/CapacitorSet/qiskit-fast-shor
\end{thebibliography}
\vspace{12pt}

>>>>

\subsection{Order-finding subroutine}\label{order-finding}
The key insight in Shor's algorithm is that the order-finding problem can be expressed in terms of finding the eigenvalues of a unitary operator, a problem which can be solved with the quantum phase estimation algorithm in $O((\log N)^3)$, i.e. linearly in the number of bits of $N$. For comparison, the best classical algorithm for factoring has a complexity of $O(e^{(\log N)^{1/3}(\log\log N)^{2/3}})$.

Shor exploits the unitary operator
\begin{equation}
U_a: x \rightarrow ax \mod N
\end{equation}
whose eigenvalues\footnote{The reader can prove this by substitution: the eigenvectors take the form $\frac{1}{\sqrt{r}}\sum_{m=0}^{r-1}\exp(\frac{-2\pi i m n}{r})\ket{x^k \mod N}$.} are a function of $r$:
\begin{equation}
e^{2i\pi k/r}\text{ for some integer }k
\end{equation}

@@

\subsection{Modular exponentiation}
The modular exponentiation procedure is at the heart of Shor's algorithm, and can be implemented in several ways. Notably, any classical computation (thus including modular exponentiation) can be expressed in terms of Toffoli gates, as they are universal for Boolean functions --- for an example of such a general approach, see \cite{toffoli}. However, in literature we can find specialized circuits for modular exponentiation, the first being described in a paper \cite{vbe} by Vedral, Barenco and Ekert and still used today in Qiskit's implementation of Shor's algorithm.

@@

\subsection{Application}
We develop a simple client-server application that emulates an email service. Users can log in with their username and password, and once logged in they can browse their email messages. It uses a simple cryptographic protocol that simulates TLS, where the server sends its public key to the client, which in turn sends an encrypted AES key (a "session key") allowing for secure two-way communication. The goal of this demonstration is to capture the public key and factor it, allowing a third party to decrypt messages intended for the server and snoop on the credentials and the email messages.

@@

\begin{thebibliography}{00}
\bibitem{shor} P. W. Shor, "Algorithms for quantum computation: discrete logarithms and factoring", SIAM J.Sci.Statist.Comput. 26
\bibitem{rsa} R. Rivest, A. Shamir, L. Adleman, "A Method for Obtaining Digital Signatures and Public-Key Cryptosystems", Communications of the ACM 21
\bibitem{miller} G. L. Miller, "Riemann's hypothesis and tests for primality", J. Comput. System Sci.
\bibitem{vbe} V. Vedral, A. Barenco, A. Ekert, "Quantum Networks for Elementary Arithmetic Operations", Physical Review A 54
\bibitem{toffoli} D. Maslov, G. W. Dueck, D. M. Miller, "Toffoli network synthesis with templates," IEEE Transactions on Computer-Aided Design of Integrated Circuits and Systems
\bibitem{ibm} IBM, "Shor's algorithm",\\https://quantum-computing.ibm.com/docs/guide/q-algos/shor-s-algorithm
\bibitem{github-repo} G. Muscarello, https://github.com/CapacitorSet/qiskit-fast-shor
\end{thebibliography}
\vspace{12pt}
